It was decided to keep the information as it is in BOLD, unless the data was missing and it was available in GENBANK
```{r echo=FALSE, warning=FALSE, message=FALSE}
library(readr)
genbank<-read.csv("../data/genbank/infotax_ncbi.csv")
bold<-read.csv("../data/bold/taxinfo_bold.csv")
#change genbank column names to match the data from genbank, to make future merging easier
colnames(genbank)<-c("Tax_id","kingdom", colnames(bold[,2:8]))

```



```{r echo=FALSE, warning=FALSE, message=FALSE}

bold_in_gb<-bold$species_name %in% genbank$species_name

#separate bold data into data with species matches to downloaded genbank data and no species matches to gb data 

#species in bold that match to gb
bold_ingb<-bold[bold_in_gb==T,]
#species in bold with no matches to gb
bold_nogb<-bold[bold_in_gb==F,]
#list of species names with no matches to gb data
sp_miss_bold<-na.omit(bold_nogb$species_name)

write_csv(as.data.frame(na.omit(sp_miss_bold)), "../data/bold/species_miss_bold.txt", col_names = F)
rm(sp_miss_bold)
```


The list was uploaded to the Taxonomy NCBI website to obtain their tax ids 
```{r echo=FALSE, warning=FALSE, message=FALSE}
library(stringr)
sp_bold_miss_id<-read_delim("../data/bold/tax_bold_gb_id.txt", delim = "|\t" )

#make tax_id a numeric variable
sp_bold_miss_id$`	taxid`<- as.numeric(sp_bold_miss_id$`	taxid`)
sp_bold_miss_id$`	name	`<-str_remove_all(sp_bold_miss_id$`	name	`, "\t")

#From our original unique bold data, once removing those with species name = NA , and after looking them up in Taxonomy NCBI 2079 tax_ids were recovered

#From bold_nogb remove phylum id 
bold_nogb<-bold_nogb[,-9]
#Add retrieved species Ids to bold_nogb data frame 
bold_nogb$Tax_id<-NA
#temporarily split bold_nogb into two data frames one with species_name==NA the other with species names
bold_no_gb_sp_na<-bold_nogb[is.na(bold_nogb$species_name),]
bold_no_gb_sp<-bold_nogb[!is.na(bold_nogb$species_name),]

bold_no_gb_sp$Tax_id<-sp_bold_miss_id[match(bold_no_gb_sp$species_name, sp_bold_miss_id$`	name	`),4][[1]]

bold_nogb<-rbind(bold_no_gb_sp, bold_no_gb_sp_na)
rm(bold_no_gb_sp, bold_no_gb_sp_na, sp_bold_miss_id)

```

Divide genbank data into data that is also in bold and data that is unique to genbank.

```{r echo=FALSE, warning=FALSE, message=FALSE}
gb_in_bold<- genbank$species_name %in% bold$species_name

#genbank data with species matches in bold
gb_inbold<-genbank[gb_in_bold==T,]

#genbank data with no species matches in bold
gb_nobold<-genbank[gb_in_bold==F,]

```


Take data frame gb_nobold (data with no species name matches to bold data).
From this point forward until otherwise noted the data starting point is genbank entries with no species match to our BOLD data base. 

```{r echo=FALSE, warning=FALSE}
#GENUS
#Separate data between those with genus matches to bold and those without
gb_genus_bold<- gb_nobold$genus_name %in% bold$genus_name
gb_genus_inbold<-gb_nobold[gb_genus_bold==T,]
gb_genus_nobold<-gb_nobold[gb_genus_bold==F,]

gb_genus_nobold<-rbind(gb_genus_nobold, gb_genus_inbold[is.na(gb_genus_inbold$genus_name),])
gb_genus_inbold<-gb_genus_inbold[!is.na(gb_genus_inbold$genus_name),]

#To the data that has genus matching those in bold, add taxonomic information from bold for all columns (excluding tax id and species name).
gb_genus<-cbind(bold[match(gb_genus_inbold$genus_name, bold$genus_name ),1:7], gb_genus_inbold[,c(9,1)])

#SUBFAMILY
#Separate data from gb_genus_nobold between those with matches by subfamily to bold data, and those without

  gb_subfamily_bold<- gb_genus_nobold$subfamily_name %in% bold$subfamily_name
  gb_subfamily_inbold<-gb_genus_nobold[gb_subfamily_bold==T,]
  gb_subfamily_nobold<-gb_genus_nobold[gb_subfamily_bold==F,]
  
  gb_subfamily_nobold<-rbind(gb_subfamily_nobold, gb_subfamily_inbold[is.na(gb_subfamily_inbold$subfamily_name),])
  gb_subfamily_inbold<-gb_subfamily_inbold[!is.na(gb_subfamily_inbold$subfamily_name),]
  
#To the data that has subfamily matching those in bold, add taxonomic information from bold for all columns (excluding tax id, species name and genus).
gb_subfamily<-cbind(bold[match(gb_subfamily_inbold$subfamily_name, bold$subfamily_name ),1:6], gb_subfamily_inbold[,c(8,9,1)])

#FAMILY
#Separate data from gb_subfamily_nobold between those with matches by family to bold data, and those without
   
  gb_family_bold<- gb_subfamily_nobold$family_name %in% bold$family_name
  gb_family_inbold<-gb_subfamily_nobold[gb_family_bold==T,]
  gb_family_nobold<-gb_subfamily_nobold[gb_family_bold==F,]
  
  gb_family_nobold<-rbind(gb_family_nobold, gb_family_inbold[is.na(gb_family_inbold$family_name),])
gb_family_inbold<-gb_family_inbold[!is.na(gb_family_inbold$family_name),]
  
#To the data that has family matching those in bold, add taxonomic information from bold for all columns (excluding tax id, species name, genus and subfamily).
gb_family<-cbind(bold[match(gb_family_inbold$family_name, bold$family_name ),1:5], gb_family_inbold[,c(7,8,9,1)])

#ORDER
#Separate data from gb_family_nobold between those with matches by order to bold data, and those without
   
  gb_order_bold<- gb_family_nobold$order_name %in% bold$order_name
  gb_order_inbold<-gb_family_nobold[gb_order_bold==T,]
  gb_order_nobold<-gb_family_nobold[gb_order_bold==F,]
  
  gb_order_nobold<-rbind(gb_order_nobold, gb_order_inbold[is.na(gb_order_inbold$order_name),])
gb_order_inbold<-gb_order_inbold[!is.na(gb_order_inbold$order_name),]
  
#To the data that has order matching those in bold, add taxonomic information from bold for all columns (excluding tax id, species name, genus, subfamily and family)
gb_order<-cbind(bold[match(gb_order_inbold$order_name, bold$order_name ),1:4], gb_order_inbold[,c(6,7,8,9,1)])

#CLASS
#Separate data from gb_order_nobold between those with matches by order to bold data, and those without
  gb_class_bold<- gb_order_nobold$class_name %in% bold$class_name
  gb_class_inbold<-gb_order_nobold[gb_class_bold==T,]
  gb_class_nobold<-gb_order_nobold[gb_class_bold==F,]
  
  
   gb_class_nobold<-rbind(gb_class_nobold, gb_class_inbold[is.na(gb_class_inbold$class_name),])
gb_class_inbold<-gb_class_inbold[!is.na(gb_class_inbold$class_name),]
  
  #To the data that has class matching those in bold add taxonomic information from bold for all (excluding tax id, species name, genus, subfamily, family and order )   and Add remaining rows from gb_class_nobold to gb_class
  gb_class<-rbind (cbind(bold[match(gb_class_inbold$class_name, bold$class_name ),1:3], gb_class_inbold[,c(5,6,7,8,9,1)]), gb_class_nobold[,c(2:9,1)])
  
```

Data frames:

sp_bold_miss_id = tax_id for species unique to BOLD (ids retrieved from Taxonomy NCBI)
bold_ingb = subset of bold that contains matches in species name to genbank
bold_nogb= complement of bold_ingb
 #GENUS
gb_inbold= subset of genbank that contains matches in species name to bold
gb_nobold= complement of gb_inbold
    gb_genus_inbold=gb_nobold subset that contains matches to genus names in bold (removing those with NA        in genus column and moving them to gb_genus_nobold)
     gb_genus= data from gb_genus_inbold with taxonomic information from bold table starting with genus
    gb_genus_nobold=gb_genus_inbold complement plus, gb_genus_inbold data with NA on genus column
          #SUBFAMILY
          gb_subfamily_inbold=gb_genus_nobold subset containing matches to subfamily names in bold  (removing those with NA in genus column and moving them to gb_genus_nobold)
          gb_subfamily=data from gb_subfamily in bold with taxonomic information from bold table starting at subfamily
          gb_subfamily_nobold=gb_subfamily_inbold complement plus, gb_subfamily_inbold data with NA on subfamily column
                #FAMILY
                gb_family_inbold= gb_subfamily_inbold subset containing matches to family names in bold (removing those with NA in family column )
                gb_family=data from gb_family_inbold with taxonomic information from bold table starting at family level
                gb_family_nobold=gb_family_inbold complement plus gb_family_inbold data with NA on family column
                      #ORDER
                      gb_order_inbold=gb_family_nobold subset containing matches to order names in bold table (removing those with NA in order column)
                      gb_order=gb_order_inbold data with taxonomic information from bold table starting with order name
                      gb_order_nobold=gb_order_inbold complement plus gb_order_inbold data with NA in the order column
                             #CLASS
                                   gb_class_inbold=gb_class_nobold subset containing matches to class names in bold table (removing those with NA in class column)
                                   gb_class_nobold=gb_class_inbold complement plus data with NA in the class column
                                   gb_class=gb_class_inbold data with taxonomic information from bold table starting with class name, joined with data from gb_class_nobold
                                   
                                   
```{r, echo=F, warning=F, message=F}
#Join all the data into a single data frame that has NCBI Taxonomy species id, kingdom, and the rest of the taxonomic information was taken from the BOLD table when available otherwise it was left as is, as retrieved from NCBI.



bold_gb<-cbind( bold_ingb[,-9],
        gb_inbold[match(bold_ingb$species_name, gb_inbold$species_name ), 1])
colnames(bold_gb)<-colnames(bold_nogb)
all_species<-rbind (bold_gb, bold_nogb, gb_genus, gb_subfamily, gb_family, gb_order, gb_class)                    
                  
rm(bold_ingb, bold_nogb, gb_inbold, gb_genus, gb_genus_inbold, gb_genus_nobold, gb_subfamily, gb_subfamily_inbold, gb_subfamily_nobold, gb_family, gb_family_inbold, gb_family_nobold, gb_order, gb_order_inbold, gb_order_nobold, gb_class, gb_class_inbold, gb_class_nobold)

#check for rows with missing kingdoms by phylum
nk<-unique(all_species[which(is.na(all_species$kingdom)), 2])
all_species[which(all_species$phylum_name=="Glomeromycota"),1]<-"Fungi"
all_species[which(all_species$phylum_name=="Chlorarachniophyta"),1]<-"Protista"
all_species[which(all_species$phylum_name=="Cryptophycophyta"),1]<-"Chromista"
all_species[which(all_species$phylum_name=="Virus"),1]<-"Virus"
all_species[which(all_species$phylum_name=="Protista"),1]<-"Protista"
#Some data still had kingdom NA while I didnt look them up in detail they seem to be mostly viruses


#phylum with missing kingdom
# Glomeromycota = Fungi
#Chlorarachniophyta=Protista
#Cryptophycophyta=
#Virus=Virus
#Protista=Protista

###Add missing taxon
#

```


```{r, echo=F, warning=F, message=F}
write_csv(all_species, "../data/taxonomic_info_all.csv")
```
