Instructions to clean the data downloaded from Genbank

The starting point for this script is the file genbank_euk.tsv, located at the folder /data/genbank. 

1)Filter out the data that doesn't have an annotation for Mexico as country. These commands are run on shell and have to be run inside the folder /data/genbank/

```
cat genbank_euk.tsv | grep "Mexico" > genbank_euk_Mexico1.tsv
cat  genbank_euk_Mexico1.tsv | grep -v "New Mexico" > genbank_euk_Mexico2.tsv
cat genbank_euk_Mexico2.tsv | grep -v "USA" > genbank_euk_Mexico.tsv  
rm genbank_euk_Mexico1.tsv genbank_euk_Mexico2.tsv
```

Check column number in the data:

```
cat genbank_euk_Mexico.tsv | awk -F "\t" '{print NF}' | sort | uniq
```

There are discrepancies in the data regarding the number of columns, there should be 15 columns per row, with the following fields:


1) Accession version
2)Accession
3)Locus
4)Sequence length 
5)Molecule type
6)Sequence
7)Definition
8)Organism
9)Taxonomic Information
10)Organism
11)Molecule type
12)Country
13)Latitude and longitude 
14) Altitude
15)Taxonomy id

A subset of the data has an extra column, BOLD ID, which is conserved.
To fix the discrepancies, first the data was separated into three different sets, each with a different number of columns.

```
cat genbank_euk_Mexico.tsv | awk -F "\t" '{if (NF==14) print $0 }' > genbank_14.tsv 
cat genbank_euk_Mexico.tsv | awk -F "\t" '{if (NF==15) print $0 }' > genbank_15.tsv
cat genbank_euk_Mexico.tsv | awk -F "\t" '{if (NF==16) print $0 }' > genbank_16.tsv 

```

Fix the data with 14 columns:

```{r, warning=FALSE, message=FALSE}
library(readr)
gb_14 <- read_delim("../data/genbank/genbank_14.tsv", 
                            "\t", escape_double = FALSE, col_names = FALSE, 
                            trim_ws = TRUE)
gb_14_1<-gb_14[,1:5]
#Add the missing column and fill it with NA
gb_14_1$sequence<-NA
gb_1<-cbind(gb_14_1, gb_14[,6:14])
gb_1$bold_id<-NA

```

Fix the data with 16 columns:

```{r, warning=FALSE, message=FALSE}
gb_16 <- read_delim("../data/genbank/genbank_16.tsv", 
                            "\t", escape_double = FALSE, col_names = FALSE, 
                            trim_ws = TRUE)

 
bold_id<-gb_16[,15]
gb_2<-cbind(gb_16[,-15], bold_id)

```

Import data with 15 columns 

```{r, warning=FALSE, message=FALSE}
gb_15 <- read_delim("../data/genbank/genbank_15.tsv", 
                            "\t", escape_double = FALSE, col_names = FALSE, 
                            trim_ws = TRUE)
gb_15$bold_id<-NA
```
Join data.
```{r, warning=FALSE, message=FALSE}
cnames<-c("accession_version",
"accession",
"locus",
"seq_len",
"mol_type",
"sequence",
"definition",
"organism",
"taxonomy",
"source_organism",
"molecule_origin",
"country",
"lat-lon",
"altitude",
"tax_id",
"bold_id"
)
colnames(gb_1)<-cnames
colnames(gb_2)<-cnames
colnames(gb_15)<-cnames

gb<-rbind(gb_1, gb_2, gb_15)
```

```{r, warning=F, message=F}
write_csv(gb, "../data/genbank/genbank_euk.csv")
```


Case 
2175846	687269	Colletotrichum cliviicola	kingdom	Fungi	phylum	Ascomycota	class	Sordariomycetes	order	Glomerellales	family	Glomerellaceae	genus	Colletotrichum
Deje el id 2175846 porque con el otro aparecen 2 ids en vez de 1 , lo corregi  el archivo genbank1805.csv
Lo mismo paso con: 
1659837	45150	Exserohilum rostratum	kingdom	Fungi	phylum	Ascomycota	class	Dothideomycetes	order	Pleosporales	family	Pleosporaceae	genus	Exserohilum

```
 sed 's/taxon:687269/taxon:2175846/g' genbank_euk.csv > genbank_euk2.csv
 sed 's/taxon:45150/taxon:1659837/g' genbank_euk2.csv > genbank_euk.csv
 rm genbank_euk2.csv
```
