
The data downloaded from BOLD is read
```{r, error=FALSE, echo=FALSE, warning=FALSE}
bold_data <- read_delim("../data/bold/bold_data.tsv", "\t", escape_double = FALSE, trim_ws = TRUE)
bold_data<-as.data.frame(bold_data)
```

The nomenclature of some phylum was changed by hand in order to match it to the one available in NCBI.

```{r}
bold_data[bold_data$phylum_name=="Chlorarachniophyta", 10]<-"Chlorarachniophyceae"
bold_data[bold_data$phylum_name=="Glomeromycota",10]<-"Glomeromycotina"


```

First a list was generated with the phylum names:


Instructions run at the shell in the folder /data/bold

```{r}
bold_phylum<-unique(bold_data$phylum_name)
write_csv(as.data.frame(bold_phylum), "../data/bold/bold_phylum.txt", col_names = F)
```

This list was then introduced to the Taxonomy website from NCBI to obtain the phylum ids. And from the file downloaded from Taxonomy the ids alone were extracted.

```
cat tax_bold.txt | egrep -o "\|\s[0-9]+" | egrep -o "[0-9]+" > taxid.txt 

```

The ids were then introduced to entrez to perform a query and obtain the phylum name with it's kingdom

```
efetch -db taxonomy -id 6656,7711,10190,6447,2763,3398,6340,7586,5204,6231,4890,2696291,6157,6040,10232,2316800,6073,5878,3041,3208,1224,120557,214504,3195,2836,58020,1760,201174,27563,2074142,6217,29197,13809,3027,28889,183924,287987,451507,451827,451828,1264859,1312402,29178,33313,35493,33310 -format xml | xtract -pattern Taxon -element TaxId ScientificName -block "*/Taxon" -if Rank -equals kingdom -element ScientificName > phylumid_kingdom.tsv
```

Import data with phylum ids and kingdoms 
```{r, message=F, echo=FALSE, warning=FALSE}
phylumid_kingdom <- read_delim("../data/bold/phylumid_kingdom.tsv",  "\t", escape_double = FALSE, col_names = FALSE, trim_ws = TRUE)
colnames(phylumid_kingdom)<-c("phylum_id", "phylum_name", "kingdom")
```

Remove repeated phylums:
There are therefore 3 repeated phylums, looking at them manually we find 
Acanthocephala, Actinobacteria and Onycophora.

These were looked up at the Taxonomy database.
Onycophora id 2074142 is a genus from the kingdom fungi.
Onycophora id 27563 is a phylum of velvet worms.
The fungi genus was removed.
Actinobacteria id 201174 phylum.
Actinobacteria id 1760 is a class.
The class was removed.
Acanthocephala id 2316800 is a genus. 
Acanthocephala id 10232 is a phylum. 
The genus was removed. 

```{r, message=F, echo=F, warning=F}
phylumid_kingdom<-phylumid_kingdom[-(which(phylumid_kingdom$phylum_id==2074142|phylumid_kingdom$phylum_id==1760|phylumid_kingdom$phylum_id==2316800)),]
```

The phylum with NA in kingdom were looked up manually either in Taxonomy NCBI or from a different source (mostly wikipedia).


Actinobacteria       Bacteria     
Bacillariophyta      Protista     
Chlorarachniophyceae Protista
Ciliophora           Protista
Crenarchaeota        Archaea    
Thermoprotei         Archaea    
Cryptophyceae        Protista  
Foraminifera         Protista     
Ochrophyta           Protista    
Proteobacteria       Bacteria
Rhodophyta           Plantae/protista

```{r message=F, warning=F, echo=F}
phylumid_kingdom$kingdom[which(phylumid_kingdom$phylum_name=="Actinobacteria"|phylumid_kingdom$phylum_name=="Proteobacteria")]<-"Bacteria"

phylumid_kingdom$kingdom[which(phylumid_kingdom$phylum_name=="Bacillariophyta"|
                                 phylumid_kingdom$phylum_name=="Chlorarachniophyceae"|
                                phylumid_kingdom$phylum_name=="Ciliophora"|
                               phylumid_kingdom$phylum_name=="Cryptophyceae"|
                                phylumid_kingdom$phylum_name=="Ochrophyta"|
                                phylumid_kingdom$phylum_name=="Foraminifera") ]<-"Protista"

phylumid_kingdom$kingdom[which(phylumid_kingdom$phylum_name=="Rhodophyta")]<-"Plantae/protista"

phylumid_kingdom$kingdom[which(phylumid_kingdom$phylum_name=="Crenarchaeota"|phylumid_kingdom$phylum_name=="Thermoprotei")]<-"Archaea"
```


For some reason when importing our list of phylum some of the original we had were left out, the list of phylum in the original data and the one imported from NCBI are compared, the missing phylum are then searched manually and added: 
#check the part were these phyla are matched to find missing phyla
```{r, echo=F, warning=F, message=F}
missing_phylum<-bold_phylum[which(bold_phylum %in% phylum_ncbi==F)]
```
Zygomycota: Fungi
Magnoliophyta: it appears as Magnoliopsida in NCBI. *Change Magnoliopsida from the genbank files to Magnoliophyta. Viridiplantae. 
Pteridophyta: Viridiplantae
Pinophyta: Viridiplantae
Phoroniformea: It appears in NCBI as phoronida (ahorita se va a corregir)
Heterokontophyta: Protista
Cryptophyta:Protista
Pyrrophycophyta: Protista
Marine: NA (probably an environmental sample)

The missing phylum are added:
```{r}

missingphylum<-rbind( c(NA, "Zygomycota", "Fungi"),
                       c(NA, "Pteridophyta", "Viridiplantae"),
                       c(NA, "Magnoliophyta", "Viridiplantae"),
                       c(NA, "Pinophyta", "Viridiplantae"),
                      c(NA, "Heterokontophyta", "Protista"),
                      c(NA, "Cryptophyta", "Protista"),
                      c(NA, "Pyrrophycophyta", "Protista"),
                       c(NA, "Marine", NA))
colnames(missingphylum)<-c("phylum_id", "phylum_name", "kingdom")

  
  
phylumid_kingdom<-rbind(phylumid_kingdom, missingphylum)



```



Add column with kingdom to bold data and writes new file with column kingdom, as well as file with taxonomic information from bold.
```{r message=F, echo=F, warning=F}
#get positions of matching phylum
matched_phylum<-match(bold_data$phylum_name, phylumid_kingdom$phylum_name)
bold_data$kingdom<-phylumid_kingdom$kingdom[matched_phylum]
taxinfo_bold<-unique(bold_data[c(81,10,12,14,16,18,20,22)])

matched_id<-match(taxinfo_bold$phylum_name, phylumid_kingdom$phylum_name)
taxinfo_bold$phylum_ncbi_id<-phylumid_kingdom$phylum_id[matched_id]
write_csv(taxinfo_bold, "../data/bold/taxinfo_bold.csv")
```

```{r message=F, echo=F, warning=F}

write_csv(bold_data, "../data/bold/bold_data.csv" )

```

