
---
title: "EDA BOLD DATA"
output: rmarkdown::github_document
---
```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyr)
library(magrittr)
library(dplyr)
library(readr)

#Obtain number of sequences per kingdom and per phylum
options(tibble.print_max = Inf)

bold<-read.csv("../data/bold/bold_complete.csv")
```
Number of sequences per kingdom:
```{r, echo=FALSE, error=FALSE, message=FALSE}

king_phyl<-bold%>% group_by(phylum_name, kingdom) %>% summarise(n())
colnames(king_phyl)<-c("phylum", "kingdom", "num_seqs") 
king_phyl<-king_phyl %>% arrange(desc(num_seqs))
king_seqs<-bold%>%group_by(kingdom)%>%summarise(n())
king_seqs[which(is.na(king_seqs$kingdom)),1]<-"NA"
king_phyl[which(is.na(king_phyl$kingdom)),2]<-"NA"
colnames(king_seqs)<-c("Kingdom", "Number_sequences")


```


```{r, echo=FALSE, error=F, message=FALSE}

bold$latlon_info<-NA
bold[which(is.na(bold$lat)|is.na(bold$lon)),75]<-FALSE
bold[which(!is.na(bold$lat)&!is.na(bold$lon)),75]<-TRUE


latlon_kin<-bold%>%group_by(kingdom, latlon_info)%>%summarise(n())
colnames(latlon_kin)<-c("kingdom", "latlon_info", "num_seqs")
latlon_kin[18,1]<-"Chromista"
latlon_kin[18,2]<-FALSE
latlon_kin[18,3]<-0
latlon_kin[which(is.na(latlon_kin$kingdom)),1]<-"NA"
latlon_kin$total<-king_seqs[match(latlon_kin$kingdom, king_seqs$Kingdom),2]
latlon_kin<-latlon_kin%>%arrange(desc(total))

bold$locality_info<-NA

bold[which( is.na(bold$region) & is.na(bold$sector) & is.na(bold$exactsite)),76]<-FALSE

bold[which(is.na(bold$locality_info)),76]<-TRUE
locality_kin<-bold%>%group_by(kingdom, locality_info)%>%summarise(n())
colnames(locality_kin)<-c("kingdom", "locality_info", "num_seqs")
locality_kin[which(is.na(locality_kin$kingdom)),1]<-"NA"
locality_kin$total<-king_seqs[match(locality_kin$kingdom, king_seqs$Kingdom),2]
locality_kin<-locality_kin%>%arrange(desc(total))

```





```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(ggplot2) 
king_phyl<-king_phyl%>% mutate(kingdom=replace_na(kingdom, "NA")) %>% mutate(phylum=replace_na(phylum, "NA")) %>% arrange(desc(num_seqs))

kingplot<-ggplot(data=king_seqs, aes(x=reorder(Kingdom, Number_sequences), y=Number_sequences))+
           geom_bar(stat="identity", fill="skyblue")+
  ggtitle("Number of sequences per kingdom") + xlab("Kingdom") + ylab("Number of sequences") +
  coord_flip() + theme(legend.position = "none") + theme_bw() +theme(plot.title=element_text(hjust = .5))


kingdont<-bold %>% group_by(kingdom) %>% count()%>% 
ungroup() %>% 
arrange(desc(n)) 

king_other<-c("Fungi
Plantae/Protista
Protista
Chromista 
Archaea 
Bacteria
NA
", sum(kingdont[3:8,2]))
kingdont<-rbind(kingdont[1:2,],king_other)

kingdont$n<-as.numeric(kingdont$n)
kingdont<-kingdont%>%mutate(percentage=round(n/sum(n),4)*100,     lab.pos=cumsum(percentage)-percentage*.5)




kingdonut<-ggplot(data=kingdont, 
       aes(x=2, y=(percentage), fill=reorder(kingdom,n)))+
      geom_bar(stat="identity")+
       coord_polar(theta="y")+
      xlim(c(1, 2.5))+
  scale_fill_brewer(palette=4) +
  ggtitle("Sequences per kingdom")+
  labs(fill="Kingdom")+
  geom_text(aes(y=lab.pos, label=paste(percentage, "%" , sep="")), col="black") +theme_void()+theme(plot.title=element_text(hjust = .5))


colnames(king_seqs)<-c("kingdom", "Number_sequences")



```

```{r kingplot, echo=FALSE, warning=FALSE, message=FALSE}

as.data.frame(king_seqs%>%arrange(desc(Number_sequences)))
kingplot
kingdonut

```

```{r, warning=F, message=F, echo=F}
options(scipen = 999)

kin_latlon_plot<-ggplot(latlon_kin, aes(factor(reorder(kingdom,total[[1]])), num_seqs, fill=latlon_info)) +
 geom_bar(stat="identity", position = "dodge") + 
  scale_fill_brewer(palette ="Pastel1") + ggtitle("Data with coordinates per Kingdom") +
  theme(axis.text.x = element_text(angle = 90)) + xlab("Kingdom") + ylab ("Number of sequences") + labs (fill="Information on coordinates") +coord_flip()+ theme_bw()+theme(plot.title = element_text(hjust = 0.5))

kin_latlon_plot2<-ggplot(latlon_kin %>% filter(kingdom != "Metazoa" && !is.na(kingdom)), aes(factor(reorder(kingdom, total[[1]])), num_seqs, fill=latlon_info)) + geom_bar(stat="identity", position = "dodge") + scale_fill_brewer(palette = "Pastel1") + ggtitle("Data with coordinates per kingdom") + theme(axis.text.x = element_text(angle = 90))+ xlab("Kingdom") + ylab ("Number of sequences")  + labs(caption="*Data with Metazoa or NA at kingdom column were excluded from this plot.", fill="Information on coordinates")+coord_flip()+ theme_bw()+ theme(plot.title = element_text(hjust = 0.5))

kin_loc_plot<-ggplot(locality_kin, aes(factor(reorder(kingdom, total[[1]])), num_seqs, fill=locality_info)) +
 geom_bar(stat="identity", position = "dodge") + 
  scale_fill_brewer(palette ="Pastel1") + ggtitle("Data with locality information per Kingdom") +
  theme(axis.text.x = element_text(angle = 90)) + xlab("Kingdom") + ylab ("Number of sequences") + labs (fill="Information on locality") +coord_flip()+ theme_bw()+theme(plot.title = element_text(hjust = 0.5))

kin_loc_plot2<-ggplot(locality_kin %>% filter(kingdom != "Metazoa" && !is.na(kingdom)), aes(factor(reorder(kingdom, total[[1]])), num_seqs, fill=locality_info)) + geom_bar(stat="identity", position = "dodge") + scale_fill_brewer(palette = "Pastel1") + ggtitle("Data with locality information per kingdom") + theme(axis.text.x = element_text(angle = 90))+ xlab("Kingdom") + ylab ("Number of sequences")  + labs(caption="*Data with Metazoa or NA at kingdom column were excluded from this plot.", fill="locality")+coord_flip()+ theme_bw()+ theme(plot.title = element_text(hjust = 0.5))




```

Information available for latitude and longitude:

```{r message=F, warning=F, echo=F}
latlon_kin[,1:3]
```

```{r kin_latlon_plot, message=FALSE, warning=FALSE, echo=FALSE}
kin_latlon_plot
```


```{r kin_latlon_plot2, message=FALSE, warning=FALSE, echo=FALSE}
kin_latlon_plot2
```

Information available for locality:

```{r message=F, warning=F, echo=F}
locality_kin[,1:3]
```

```{r kin_loc_plot, message=FALSE, warning=FALSE, echo=FALSE}
kin_loc_plot
```


```{r kin_loc_plot2, message=FALSE, warning=FALSE, echo=FALSE}
kin_loc_plot2
```

Sequences from kingdom Metazoa

```{r message=F, warning=F, error=F, echo=F}

phyl_metplot<-ggplot(data=king_phyl[which(king_phyl$kingdom=="Metazoa" & !is.na(king_phyl$phylum)),], aes(x=reorder(phylum, num_seqs), y=num_seqs, fill=reorder(phylum,n)))+
           geom_bar(stat="identity", fill="skyblue") + ggtitle("Number of sequences per phylum for Metazoa") + xlab("Phylum") + ylab ("Number of sequences")  + theme_bw() + theme(plot.title = element_text(hjust = 0.5)) +coord_flip()


phyl_metplot2<-ggplot(data=king_phyl[which(king_phyl$kingdom=="Metazoa" & king_phyl$phylum!="Arthropoda" & king_phyl$phylum!="Chordata" & !is.na(king_phyl$phylum)),], aes(x=reorder(phylum, num_seqs), y=num_seqs, fill=phylum))+
           geom_bar(stat="identity", fill="skyblue")+theme(axis.text.x=element_text(angle=90)) + ggtitle("Sequences per phylum for Metazoa") + xlab("Phylum") + ylab ("Number of sequences")  + labs(caption = "*For this plot the phyla Chordata and Arthropoda were excluded.") + coord_flip() + theme_bw() + theme(plot.title = element_text(hjust = 0.5)) + theme(plot.caption = element_text(hjust = .5))

king_phyl[which(king_phyl$kingdom=="Metazoa"), c(1,3)]
```

```{r phyl_metplot, message=F, warning=F, echo=F}
phyl_metplot
```

```{r phyl_metplot2, message=F, warning=F, echo=F}
phyl_metplot2


```



Sequences from kingdom Viridiplantae

```{r, message=F, echo=F, warning=FALSE}

phyl_plantae<-ggplot(data=king_phyl[which(king_phyl$kingdom=="Viridiplantae"),], aes(x=reorder(phylum, num_seqs), y=num_seqs))+
           geom_bar(stat="identity", fill="skyblue")+theme(axis.text.x=element_text(angle=90)) + ggtitle("Number of sequences per phylum for Viridiplantae")  + xlab("Phylum") + ylab ("Number of  sequences")  + coord_flip() + theme_bw()+ theme(plot.title = element_text(hjust = 0.5)) 

phyl_plantae2<-ggplot(data=king_phyl[which(king_phyl$kingdom=="Viridiplantae"&king_phyl$phylum!="Magnoliophyta"),], aes(x=reorder(phylum, num_seqs), y=num_seqs)) +
     geom_bar(stat="identity", fill="skyblue") + 
    theme(axis.text.x=element_text(angle=90)) + 
    ggtitle("Number of sequences per phylum for Viridiplantae") + 
    xlab("Phylum") + 
    ylab ("Number of  sequences")  + 
    coord_flip() + 
    theme_bw() + 
    theme(plot.title = element_text(hjust = 0.5)) + labs(caption="*Phylum Magnoliophyta was excluded from this plot.") + theme(plot.caption = element_text(hjust = 0.5))

king_phyl[which(king_phyl$kingdom=="Viridiplantae"),c(1,3)]
```

```{r phyl_plantae, message=F, warning=F, echo=FALSE}
phyl_plantae
```


```{r phyl_plantae2, message=F, warning=F, echo=FALSE}
phyl_plantae2
```

```{r, echo=F, message=F, warning=F}
phyl_fungi<-ggplot(data=king_phyl[which(king_phyl$kingdom=="Fungi" & !is.na(king_phyl$phylum)),], aes(x=reorder(phylum, num_seqs), y=num_seqs))+
           geom_bar(stat="identity", fill="skyblue")+theme(axis.text.x=element_text(angle=90)) + ggtitle("Sequences per phylum for Fungi") + coord_flip()+ theme_bw()+ theme(plot.title = element_text(hjust = 0.5))  + xlab("Phylum") + ylab ("Number of sequences")

phyl_fungi2<-ggplot(data=king_phyl[which(king_phyl$kingdom=="Fungi" & king_phyl$phylum!="Ascomycota" & king_phyl$phylum!="Basidiomycota" & !is.na(king_phyl$phylum)),], aes(x=reorder(phylum, num_seqs), y=num_seqs))+                   
           geom_bar(stat="identity", fill="skyblue")+theme(axis.text.x=element_text(angle=90)) + ggtitle("Sequences per phylum for Fungi") + coord_flip()+ theme_bw()+theme(plot.title = element_text(hjust = 0.5)) +labs(caption = "*Phylum Ascomycota and Basidiomycota were excluded from this plot")  + xlab("Phylum") + ylab ("Number of sequences") +theme(plot.caption = element_text(hjust = 0.5))


fun_phyl<-king_phyl[which(king_phyl$kingdom=="Fungi"),c(1,3)]
other_fun<-fun_phyl[4:12,]
fundont1<-fun_phyl[1:3,]
fundont1[4,1]<-"Other phyla"
fundont1[4,2]<-sum(other_fun$num_seqs)
fundont2<-other_fun[1:3,]
fundont2[4,1]<-paste0("\t \n",other_fun[4,1],"\t \n",  other_fun[5,1],"\t \n", other_fun[6,1], "\t \n", other_fun[7,1], "\t \n", other_fun[8,1], "\t \n", other_fun[9,1] , "\t \n", sep = " \n")
fundont2[4,2]<-sum(other_fun[4:8,2])



fun_phyl
```

```{r phyl_fungi, message=F, warning=F}
phyl_fungi
```


Sequences from kingdom Protista
```{r, echo=F, message=F, warning=F}
king_phyl[which(king_phyl$kingdom=="Protista" & !is.na(king_phyl$phylum) | king_phyl$kingdom=="Plantae/protista"),]
```

```{r, echo=F, message=F, warning=F}
phyl_prot<-ggplot(data=king_phyl[which(king_phyl$kingdom=="Protista" & !is.na(king_phyl$phylum) | king_phyl$kingdom=="Plantae/protista"),], aes(x=reorder(phylum, num_seqs), y=num_seqs))+
           geom_bar(stat="identity", fill="skyblue")+theme(axis.text.x=element_text(angle=90)) + ggtitle("Sequences per phylum for Protista") + coord_flip()+theme_bw()+ theme(plot.title = element_text(hjust = 0.5))  + xlab("Phylum") + ylab ("Number of sequences")  + labs (caption="*Phylum Rhodophyta (included here) is the only phylum classified as Plantae/Protista") + theme(plot.caption = element_text(hjust = 0.5))
phyl_prot
```

Sequences from kingdom Archaea

```{r, echo=F, message=F, warning=F}
king_phyl[king_phyl$kingdom=="Archaea",]
phyl_arc<-ggplot(data=king_phyl[which(king_phyl$kingdom=="Archaea" & !is.na(king_phyl$phylum)),], aes(x=phylum, y=num_seqs))+
           geom_bar(stat="identity", fill="skyblue") + ggtitle("Sequences per phylum for Archaea") + coord_flip() + theme_bw() + theme(plot.title = element_text(hjust = 0.5)) + xlab("Phylum") + ylab ("Number of sequences")
```

Detailed analysis into phyla of interest:

Metazoa: Chordata and Arthropoda

Viridiplantae: Magnoliophyta and Pinophyta

Fungi: Ascomycota and Basidiomycota
```{r, echo=F, message=F, warning=F}
#Filter to keep phyla we want
bold_ph<-bold %>% filter(phylum_name ==  "Chordata" | 
                                 phylum_name ==  "Arthropoda" |
                                 phylum_name ==  "Basidiomycota" |
                                 phylum_name ==  "Ascomycota" |
                                 phylum_name ==  "Magnoliophyta" |
                                 phylum_name == "Pinophyta"
                                 )

#Exclude model organisms

model_org<-c('Ashbya gossypii', 'Aspergillus nidulans', 'Coprinus cinereus', 'Cryptococcus neoformans', 'Neurospora crassa', 'Saccharomyces cerevisiae', 'Schizophyllum commune', 'Schizosaccharomyces pombe', 'Arabidopsis thaliana', 'Selaginella moellendorffii', 'Brachypodium distachyon', 'Setaria viridis', 'Lotus japonicus', 'Lemna gibba', 'Medicago truncatula', 'Erythranthe guttata', 'Mimulus guttatus', 'Nicotiana benthamiana', 'Oryza sativa', 'Physcomitrella patens', 'Marchantia polymorpha', 'Amphimedon queenslandica',  'Aplysia cornigera', 'Aplysia fasciata', 'Caenorhabditis elegans', 'Drosophila melanogaster',  'Euprymna scolopes', 'Macrostomum lignano', 'Pristionchus pacificus', 'Scathophaga stercoraria', 'Schmidtea mediterranea', 'Felis sylvestris catus', 'Canis lupus familiaris', 'Mesocricetus auratus', 'Cavia porcellus', 'Oryzias latipes', 'Mus musculus', 'Heterocephalus glaber', 'Columba livia domestica', 'Rattus norvegicus', 'Rhesus macaque', 'Xenopus laevis', 'Xenopus tropicalis', 'Danio rerio', 'Taeniopygia guttata', 'Homo sapiens')
bold_ph<-bold_ph[-(which(!(is.na(match(bold_ph$species_name, model_org))))),]




```

Analysis on genetic markers available for the selected organisms.


```{r, echo=F, message=F, warning=F}
bold_ph$marker<-bold_ph$markercode
bold_ph[which(is.na(bold_ph$marker)),77]<-bold_ph[which(is.na(bold_ph$marker)),64]
bold_ph$marker<-sub("\\|.*", "", bold_ph$marker)

#write file bold_ph which filters phyla we want and exclude model organisms and the added column that combines the marker information from columns  markercode and marker_codes
#write bold data with the column 
write.csv(bold_ph, "../data/bold/bold_ph.csv")


metazoa_markers<-bold_ph %>% filter(kingdom=="Metazoa") %>% group_by(marker) %>% summarise(n())
colnames(metazoa_markers)<-c("Marker", "Number_sequences")

metazoa_markers<-metazoa_markers%>% mutate(Marker=replace_na(Marker, "NA"))  %>% arrange(desc(Number_sequences))


metazoa_latlon_markers<-bold_ph %>% filter(kingdom=="Metazoa") %>% group_by(marker, latlon_info) %>% summarise (n())
colnames(metazoa_latlon_markers)<-c("Marker", "latlon_info", "Number_sequences")
metazoa_latlon_markers[which(is.na(metazoa_latlon_markers$Marker)),1]<-"NA"
metazoa_latlon_markers$total<-metazoa_markers[match(metazoa_latlon_markers$Marker, metazoa_markers$Marker),2]
metazoa_latlon_markers<-metazoa_latlon_markers %>% arrange(desc(total))

metazoa_loc_markers<-bold_ph %>% filter(kingdom=="Metazoa") %>% group_by(marker, locality_info) %>% summarise (n())
colnames (metazoa_loc_markers)<-c("Marker", "locality_info", "Number_sequences")
metazoa_loc_markers[which(is.na(metazoa_loc_markers$Marker)),1]<-"NA"
metazoa_loc_markers$total<-metazoa_markers[match(metazoa_loc_markers$Marker, metazoa_markers$Marker),2]
metazoa_loc_markers<-metazoa_loc_markers %>% arrange(desc(total))


```

Genetic markers available for all organisms:
```{r, echo=F, message=F, warning=F}
bold_ph%>%group_by(marker)%>% summarise(n())%>%arrange(desc(`n()`))

```
Number of sequences per marker for Metazoa
```{r, echo=F, message=F, warning=F}

metazoa_latlon_markers[,1:3]
metazoa_loc_markers[,1:3]
met_marker_plot<-ggplot(data=metazoa_markers, aes(x=reorder(Marker, Number_sequences), y=Number_sequences))+
           geom_bar(stat="identity", fill="skyblue")+
  ggtitle("Number of sequences per marker for Metazoa") + xlab("Marker") + ylab("Number of sequences") +
  coord_flip() + theme(legend.position = "none") + theme_bw() +theme(plot.title=element_text(hjust = .5)) + labs (caption="*Phyla included: Chordata and Arthropoda") + theme(plot.caption = element_text(hjust = 0.5))
met_marker_plot

metmarker_latlon_plot<-ggplot(metazoa_latlon_markers[1:20,], aes(factor(reorder(Marker,total[[1]])), Number_sequences, fill=latlon_info)) +
 geom_bar(stat="identity", position = "dodge") + 
  scale_fill_brewer(palette ="Pastel1") + ggtitle("Data with coordinates per marker for Metazoa") +
  theme(axis.text.x = element_text(angle = 90)) + xlab("Marker") + ylab ("Number of sequences") + labs (fill="Information on coordinates", caption = "*Only the ten most abundant markers are included in this plot") +coord_flip()+ theme_bw()+theme(plot.title = element_text(hjust = 0.5)) + theme(plot.caption = element_text(hjust = 0.5))
metmarker_latlon_plot


metmarker_loc_plot<-ggplot(metazoa_loc_markers[1:20,], aes(factor(reorder(Marker, total[[1]])), Number_sequences, fill=locality_info)) + geom_bar(stat="identity", position = "dodge") + scale_fill_brewer(palette = "Pastel1") + ggtitle("Data with locality information per marker for Metazoa") + theme(axis.text.x = element_text(angle = 90))+ xlab("Marker") + ylab ("Number of sequences")  + labs(fill="Information on locality", caption = "*Only the ten most abundant markers are included in this plot")+coord_flip()+ theme_bw()+ theme(plot.title = element_text(hjust = 0.5)) + theme(plot.caption = element_text(hjust = 0.5))
metmarker_loc_plot


```

Number of sequences per marker for Viridiplantae
```{r, echo=F, message=F, warning=F}

viridiplantae_markers<-bold_ph %>% filter(kingdom=="Viridiplantae") %>% group_by(marker) %>% summarise(n())
colnames(viridiplantae_markers)<-c("Marker", "Number_sequences")

viridiplantae_markers<-viridiplantae_markers%>% mutate(Marker=replace_na(Marker, "NA"))  %>% arrange(desc(Number_sequences))


viridiplantae_latlon_markers<-bold_ph %>% filter(kingdom=="Viridiplantae") %>% group_by(marker, latlon_info) %>% summarise (n())
colnames(viridiplantae_latlon_markers)<-c("Marker", "latlon_info", "Number_sequences")
viridiplantae_latlon_markers[which(is.na(viridiplantae_latlon_markers$Marker)),1]<-"NA"
viridiplantae_latlon_markers$total<-viridiplantae_markers[match(viridiplantae_latlon_markers$Marker, viridiplantae_markers$Marker),2]
viridiplantae_latlon_markers<-viridiplantae_latlon_markers %>% arrange(desc(total))

viridiplantae_loc_markers<-bold_ph %>% filter(kingdom=="Viridiplantae") %>% group_by(marker, locality_info) %>% summarise (n())
colnames (viridiplantae_loc_markers)<-c("Marker", "locality_info", "Number_sequences")
viridiplantae_loc_markers[which(is.na(viridiplantae_loc_markers$Marker)),1]<-"NA"
viridiplantae_loc_markers$total<-viridiplantae_markers[match(viridiplantae_loc_markers$Marker, viridiplantae_markers$Marker),2]
viridiplantae_loc_markers<-viridiplantae_loc_markers %>% arrange(desc(total))


```

```{r, echo=F, message=F, warning=F}

viridiplantae_latlon_markers[,1:3]
viridiplantae_loc_markers[,1:3]
virp_marker_plot<-ggplot(data=viridiplantae_markers, aes(x=reorder(Marker, Number_sequences), y=Number_sequences))+
           geom_bar(stat="identity", fill="skyblue")+
  ggtitle("Number of sequences per marker for Viridiplantae") + xlab("Marker") + ylab("Number of sequences") +
  coord_flip() + theme(legend.position = "none") + theme_bw() +theme(plot.title=element_text(hjust = .5)) + labs (caption="*Phyla included: Pinophyta and Magnoliophyta") + theme(plot.caption = element_text(hjust = 0.5))
virp_marker_plot

virpmarker_latlon_plot<-ggplot(viridiplantae_latlon_markers[,], aes(factor(reorder(Marker,total[[1]])), Number_sequences, fill=latlon_info)) +
 geom_bar(stat="identity", position = "dodge") + 
  scale_fill_brewer(palette ="Pastel1") + ggtitle("Data with coordinates per marker for Viridiplantae") +
  theme(axis.text.x = element_text(angle = 90)) + xlab("Marker") + ylab ("Number of sequences") + labs (fill="Information on coordinates", caption = "") +coord_flip()+ theme_bw()+theme(plot.title = element_text(hjust = 0.5)) + theme(plot.caption = element_text(hjust = 0.5))
virpmarker_latlon_plot


virpmarker_loc_plot<-ggplot(viridiplantae_loc_markers[,], aes(factor(reorder(Marker, total[[1]])), Number_sequences, fill=locality_info)) + geom_bar(stat="identity", position = "dodge") + scale_fill_brewer(palette = "Pastel1") + ggtitle("Data with locality information per marker for Viridiplantae") + theme(axis.text.x = element_text(angle = 90))+ xlab("Marker") + ylab ("Number of sequences")  + labs(fill="Information on locality", caption = "")+coord_flip()+ theme_bw()+ theme(plot.title = element_text(hjust = 0.5)) + theme(plot.caption = element_text(hjust = 0.5))
virpmarker_loc_plot


```

```{r, echo=F, message=F, warning=F}

fungi_markers<-bold_ph %>% filter(kingdom=="Fungi") %>% group_by(marker) %>% summarise(n())
colnames(fungi_markers)<-c("Marker", "Number_sequences")

fungi_markers<-fungi_markers%>% mutate(Marker=replace_na(Marker, "NA"))  %>% arrange(desc(Number_sequences))


fungi_latlon_markers<-bold_ph %>% filter(kingdom=="Fungi") %>% group_by(marker, latlon_info) %>% summarise (n())
colnames(fungi_latlon_markers)<-c("Marker", "latlon_info", "Number_sequences")
fungi_latlon_markers2<-as.data.frame(rbind(c("5.8S", FALSE, 0), c("ITS1", FALSE, 0), c("ITS2", FALSE, 0), c("28S", TRUE, 0), c("NA", TRUE, 0)))
fungi_latlon_markers2$V2<-as.logical(fungi_latlon_markers2$V2)
fungi_latlon_markers2$V3<-as.numeric(fungi_latlon_markers2$V3)
colnames(fungi_latlon_markers2)<-c("Marker", "latlon_info", "Number_sequences")
fungi_latlon_markers<-rbind(fungi_latlon_markers, fungi_latlon_markers2)
fungi_latlon_markers[which(is.na(fungi_latlon_markers$Marker)),1]<-"NA"
fungi_latlon_markers$total<-fungi_markers[match(fungi_latlon_markers$Marker, fungi_markers$Marker),2]
fungi_latlon_markers<-fungi_latlon_markers %>% arrange(desc(Marker)) %>% arrange(desc(total))







fungi_loc_markers<-bold_ph %>% filter(kingdom=="Fungi") %>% group_by(marker, locality_info) %>% summarise (n())
colnames (fungi_loc_markers)<-c("Marker", "locality_info", "Number_sequences")
fungi_loc_markers2<-as.data.frame(rbind(c("28S", FALSE, 0),c("NA", FALSE, 0)))
colnames(fungi_loc_markers2)<- c("Marker", "locality_info", "Number_sequences")
fungi_loc_markers2$locality_info<-as.logical(fungi_loc_markers2$locality_info)
fungi_loc_markers2$Number_sequences<-as.numeric(fungi_loc_markers2$Number_sequences)
fungi_loc_markers[which(is.na(fungi_loc_markers$Marker)),1]<-"NA"
fungi_loc_markers<-rbind(fungi_loc_markers2, fungi_loc_markers)
fungi_loc_markers$total<-fungi_markers[match(fungi_loc_markers$Marker, fungi_markers$Marker),2]
fungi_loc_markers<-fungi_loc_markers %>% arrange(desc(Marker))%>% arrange(desc(total))


```
Number of sequences per marker for Fungi
```{r, echo=F, message=F, warning=F}

fungi_latlon_markers[,1:3]
fungi_loc_markers[,1:3]
fungi_marker_plot<-ggplot(data=fungi_markers, aes(x=reorder(Marker, Number_sequences), y=Number_sequences))+
           geom_bar(stat="identity", fill="skyblue")+
  ggtitle("Number of sequences per marker for Fungi") + xlab("Marker") + ylab("Number of sequences") +
  coord_flip() + theme(legend.position = "none") + theme_bw() +theme(plot.title=element_text(hjust = .5)) + labs (caption="*Phyla included: Ascomycota and Basidiomycota") + theme(plot.caption = element_text(hjust = 0.5))
fungi_marker_plot

fungimarker_latlon_plot<-ggplot(fungi_latlon_markers[,], aes(factor(reorder(Marker,total[[1]])), Number_sequences, fill=latlon_info)) +
 geom_bar(stat="identity", position = "dodge") + 
  scale_fill_brewer(palette ="Pastel1") + ggtitle("Data with coordinates per marker for Fungi") +
  theme(axis.text.x = element_text(angle = 90)) + xlab("Marker") + ylab ("Number of sequences") + labs (fill="Information on coordinates", caption = "") +coord_flip()+ theme_bw()+theme(plot.title = element_text(hjust = 0.5)) + theme(plot.caption = element_text(hjust = 0.5))
fungimarker_latlon_plot


fungimarker_loc_plot<-ggplot(fungi_loc_markers[,], aes(factor(reorder(Marker, total[[1]])), Number_sequences, fill=locality_info)) + geom_bar(stat="identity", position = "dodge") + scale_fill_brewer(palette = "Pastel1") + ggtitle("Data with locality information per marker for Fungi") + theme(axis.text.x = element_text(angle = 90))+ xlab("Marker") + ylab ("Number of sequences")  + labs(fill="Information on locality", caption = "")+coord_flip()+ theme_bw()+ theme(plot.title = element_text(hjust = 0.5)) + theme(plot.caption = element_text(hjust = 0.5))
fungimarker_loc_plot


```
